---
title: "NOAA Weather Stations Map using leaflet"
author: "Mark S Spoto"
date: "August 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
packages <- c("lubridate", "dplyr", "ggplot2", "RColorBrewer", "mice", "VIM", "grid", "gridExtra","leaflet")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org")  
}

library(leaflet)
library(dplyr)
```

## Introduction

This presentation will show all the NOAA Weather Stations througout the United States using leaflet.  Below is a list of weather
stations from ["https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/ghcnd-stations.txt"].  Below are the file definitions.

Variable  | Columns | Type
--------- | ------- | ---------  
ID        | 1-11    | Character  
LATITUDE  | 13-20   | Real  
LONGITUDE | 22-30   | Real  
ELEVATION | 32-37   | Real  
STATE     | 39-40   | Character  
NAME      | 42-71   | Character  
GSN FLAG  | 73-75   | Character  
HCN/CRN FLAG | 77-79 |Character  
WMO ID       | 81-85 | Character

```{r, echo=FALSE}
totalUSStations <- 0
con = file("./stations.txt", "r")
data <- readLines(con)
for (i in 1:length(data)) {
  if (startsWith(data[i], "US")) {
    totalUSStations <- totalUSStations + 1
  }
}
close(con)
```

### Key Notes about the Map
Here are some lessons learned about using leaflet to show maps.  The dataset contains over 100 thousand locations througout the world.  This presentation will only use the stations found in the United States which is approximately `r format(totalUSStations, scientific=FALSE, big.mark=",")`.

Due to the large number of records, the clusterOptions will be used to render the page faster.  The loading time for the dataset takes approximately 2 to 3 minutes to load the dataframe.  Creating the dataframe in memory before loading the data was the fastest time to loading the data into the dataframe.

## NOAA Weather Stations in the United States
Click on the clusters to view the pins.  Each pin will display the state and station name.  To see an io slide version click on this link: "https://markspoto.github.io/NOAAStationsMarkdown/NOAAStationsSlides.Rmd""

```{r echo=FALSE, warning=FALSE}
trim <- function(x) gsub("^\\s+|\\s+$", "", x)

n <- totalUSStations
idx <- 1
df = data.frame(id = character(n), lat = numeric(n), lng = numeric(n), st = character(n), name = character(n), desc = character(n), stringsAsFactors = FALSE)
con = file("./stations.txt", "r")
data <- readLines(con)
for (i in 1:length(data)) {
  if (startsWith(data[i], "US")) {
    s = data[i]
    df$id[idx] <- substr(s, 1, 11)
    df$lat[idx] <- as.numeric(substr(s, 13, 20))
    df$lng[idx] <- as.numeric(substr(s, 22, 30))
    df$st[idx] <- substr(s, 39, 40)
    df$name[idx] <- trim(substr(s, 42, 71))
    df$desc[idx] <- paste("State:", df$st[idx], " Station:", df$name[idx])
    idx = idx + 1
  }
}
close(con)

stationpopup <- df[,c("desc")]
df[,c("lat","lng")] %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(popup = stationpopup, clusterOptions = markerClusterOptions())
```





